---
title: "Regression Analysis Assignment"
author: "Sercan Doğan"
date: "October 12, 2017"
output:
  html_document: default
  pdf_document: default
---


A) 
- Estimate the model $\hat{Y} = \hat{\beta_0} + \hat{\beta_1}X$  
- Investigate that your model will be valid for further analysis such as statistical inference, prediction and estimation, confidence intervals
- Investigate behaviour of standardized residuals.

B)
- Estimate the model $\hat{Y} = \hat{\beta_0} + \hat{\beta_1}X + \hat{\beta_2}X^2$ (polynomial regression)
- Investigate residuals using graphic.

Which model is more suitable to data at hand?


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```


```{r}
rdata <- tribble(
  ~x, ~y,
  1,  23,
  2,  29,
  3,  49,
  4,  64,
  4,  74,
  5,  87,
  6,  96,
  6,  97,
  7,  109,
  8,  119,
  9,  149,
  9,  145,
  10, 154,
  10, 166
)

rdata
```


### A

For first model we need to assign `rdata` to new tibble. Because we will create specific columns for each model. 

```{r}
model_a <- rdata
```

Total observation count:
```{r}
n <- nrow(model_a)
n
```


Let's see scatter plot of data:
```{r}
ggplot(model_a, aes(x = x, y = y)) +
  geom_point()

```

Means
```{r}
mean_x <- sum(model_a$x) / nrow(model_a)
mean_y <- sum(model_a$y) / nrow(model_a)
#X
mean_x
#Y
mean_y

```

We will use formula which you see below for determinating $\hat{\beta_0}$ and $\hat{\beta_1}$

$$\hat{\beta_1} = \dfrac{S_{xy}}{S_{xx}} = \dfrac{\sum{XY} - n\bar{X}\bar{Y}}{\sum{X^2} - n\bar{X}^2}$$
$$\hat{\beta_0} = \bar{Y} - \hat{\beta_1}\bar{X}$$

So, we need $X^2$ and $XY$:

```{r}
model_a <- model_a %>%
  mutate(xsquare = x ^ 2,
         xy = x * y)

model_a

```
Thus,

```{r}
slope <- (sum(model_a$xy) - (n * mean_x * mean_y)) / (sum(model_a$xsquare) - (n * mean_x ^ 2))


intercept <- mean_y - slope * mean_x

paramaters <- list('intercept' = intercept, 'slope' = slope)

paramaters

```


Determine residual:

```{r}
model_a <- model_a %>% 
  mutate(y_estimator = paramaters$intercept + x * paramaters$slope,
         e = y - y_estimator,
         esquare = e ^ 2)

model_a

```


Sum of residual must be 0;
```{r}
round(sum(model_a$e),2)
```

Mean Square Error:
```{r}
mse <- sum(model_a$esquare / (n - 2))

mse
```

Add standardized residual as a column:
```{r}
model_a <- model_a %>%
  mutate(estd = e / sqrt(mse))


model_a$estd
```


Random pattern appears so it is valid for further analysis and the form of the estimated model is appropriate.
```{r}
ggplot(model_a,aes(x,estd)) +
  geom_point()

```

Let's draw regression line.

```{r}
ggplot(model_a, mapping = aes(x,y)) + 
  geom_point() + geom_line(aes(y=model_a$y_estimator))

```





### B
First of all;

```{r}
model_b <- rdata

model_b
```


Our polynomial model is $\hat{Y} = \hat{\beta_0} + \hat{\beta_1}X + \hat{\beta_2}X^2$



$$e_i = Y_i - \hat{Y_i} = Y_i - \hat{\beta_0} - \hat{\beta_1}X_i- \hat{\beta_2}X_i^2 $$
And then,

$$S = \sum{e_i^2} = \sum{(Y_i - \hat{Y_i})^2} = \sum{(Y_i - \hat{\beta_0} - \hat{\beta_1}X - \hat{\beta_2}X^2)^2}$$

When we derivate of $\beta_0$, $\beta_1$, $\beta_2$ and then equal to zero, we get three system of equations:

$$\sum{Y_i} = n\hat{\beta_0} + \hat{\beta_1}\sum{X_i} + \hat{\beta_2}\sum{X_i^2}$$
$$\sum{X_iY_i} = \hat{\beta_0}\sum{X_i} + \hat{\beta_1}\sum{X_i^2} + \hat{\beta_2}\sum{X_i^3}$$
$$\sum{X_i^2Y_i} = \hat{\beta_0}\sum{X_i^2} + \hat{\beta_1}\sum{X_i^3} + \hat{\beta_2}\sum{X_i^4}$$
Thus,

We create matrix of cofficients and equations vector. 

```{r}
system_equation <- matrix(c(n, sum(model_b$x), sum(model_b$x ^ 2),
                            sum(model_b$x), sum(model_b$x ^ 2), sum(model_b$x ^ 3),
                            sum(model_b$x ^ 2), sum(model_b$x ^ 3), sum(model_b$x ^4)),
                          ncol = 3,
                          byrow = T)

equal_to <- c(sum(model_b$y),
              sum(model_b$x * model_b$y), 
              sum(model_b$x ^ 2 * model_b$y))


coefficients_b <- solve(system_equation,equal_to)

names(coefficients_b) <- c("beta0", "beta1", "beta2")

coefficients_b <- as.list(coefficients_b)

coefficients_b


```

```{r}
model_b <- model_b %>% 
  mutate(y_estimator = coefficients_b$beta0 + x * coefficients_b$beta1 + coefficients_b$beta2 * x ^ 2,
         e = y - y_estimator,
         esquare = e ^ 2)

model_b
```


Sum of residual must be 0;
```{r}
round(sum(model_b$e),2)
```


Mean Square Error:
```{r}
mse <- sum(model_b$esquare / (n - 3))

mse
```

Add standardized residual as a column:
```{r}
model_b <- model_b %>%
  mutate(estd = e / sqrt(mse))


model_b$estd
```



Random pattern appears:
```{r}
ggplot(model_b,aes(x,estd)) +
  geom_point()

```


Regression line:
```{r}



ggplot(model_b, mapping = aes(x,y)) + 
  geom_point() + geom_line(aes(y=model_b$y_estimator))
```


Finding the best model is to find a model which makes minimum of sum of square of residual

```{r}
# arg min --> sum E^2
sum(model_a$esquare) 


sum(model_b$esquare)

```


So, Model b which is Polynomial is more suitable.

